---
title: "Folder Tree"
author: "Derek Sollberger"
format: html
---

I saw some tweet where R programmers were asking about how to visualize folders and subfolders among their work files, so I thought I would try it out too.  One of the answers was these Stack Overflow threads:

* [https://stackoverflow.com/questions/67357402/how-to-i-see-visual-of-my-folder-structure-in-rstudio](https://stackoverflow.com/questions/67357402/how-to-i-see-visual-of-my-folder-structure-in-rstudio)
* [https://stackoverflow.com/questions/36094183/how-to-build-a-dendrogram-from-a-directory-tree](https://stackoverflow.com/questions/36094183/how-to-build-a-dendrogram-from-a-directory-tree)

```{r}
library("data.tree")
library("dplyr")
```


First, I set the working directory to where I have most of my GitHub-linked files.  Next, we collect the paths (along with subdirectories)

```{r}
paths = unique(c(list.dirs(full.names = T),list.files(full.names = T)))
```

It appears that the information is collected into a data frame.

```{r}
x <- lapply(strsplit(paths, "/"), function(z) as.data.frame(t(z)))
x <- plyr::rbind.fill(x) #plyr??
```

To restore individual file names, we `collapse` the information with the slashes.

```{r}
x$pathString <- apply(x, 1, function(x) paste(trimws(na.omit(x)), collapse="/"))
```

Now, we can build the tree.

```{r}
mytree <- data.tree::as.Node(x)
```

Do we dare `print` the tree?

```{r}
print(mytree)
```

