---
title: "Making vdist"
author: "Derek Sollberger"
date: "2025-12-01"
format:
  html:
    toc: true
---

# Goal

Here, I hope to document the process(es) to make the `vdist` R package.  This package will make `ggplot2` wrappers to ease instruction of elementary probability distributions.

We are motivated by the December 2025 LinkedIn posts by [Dr Athanasia Monika Mowinckel](https://www.linkedin.com/in/drmowinckels/). She is guiding us readers in R package development over the course of a [25-day journey](https://www.linkedin.com/feed/update/urn:li:activity:7400837138604290048/)


## Day 1: usethis

For [Day 1](https://www.linkedin.com/feed/update/urn:li:activity:7401213859547652096/), Dr Monwinckel advises us to enjoy the ease of using the [usethis](https://usethis.r-lib.org/) package.  For my `vdist` package, I am placing the files (within a Windows operating system) into the `Documents\R` folder.

```{r}
#| eval: false

# create space for new package
# (opens new instance of RStudio)
usethis::create_package("vdist")

# set usage license
usethis::use_mit_license()

# create README as R markdown document
# (opens new .rmd file)
usethis::use_readme_rmd()

# setup GitHub infrastructure with "standard" workflow
usethis::use_github_action("check-standard")

# setup testthat infrastructure ("3" will be explained later?)
usethis::use_testthat(3)

# setup package-down infrastructure 
# (opens new .yml file)
usethis::use_pkgdown()
```

Furthermore, we were advised to edit our RProfile page with

```{r}
#| eval: false

usethis::edit_r_profile()
```

and then adding to the RProfile file

```
options(usethis.full_name = "Derek Sollberger")
options(usethis.protocol = "ssh")
```











::: {.callout-note collapse="true"}
## Session Info

```{r}
sessionInfo()
```
:::


:::: {.columns}

::: {.column width="45%"}
	
:::

::: {.column width="10%"}
	
:::

::: {.column width="45%"}

:::

::::

::::: {.panel-tabset}



:::::