---
title: "Highcharter"
author: "Derek Sollberger"
date: "2024-02-21"
format: html
---

Recently, a colleague told us about the `highcharter` data visualization package.  It seems to have wrappers for Javascript functionality, so we could aim for pop-up information and other interactivity.  Perhaps `highcharter` is more robust than `plotly` in some ways, but we cannot simply add to `ggplot2` code has `highcharter` has its own syntax.

```{r}
#| message: false
#| warning: false
library("highcharter")
library("palmerpenguins")
library("tidyverse")
```

## Grammer of Graphics

We will create a simple scatterplot in `ggplot2`.

```{r}
#| message: false
#| warning: false
penguins |>
  ggplot() +
  geom_point(aes(x = flipper_length_mm, y = bill_length_mm,
             color = species)) +
  labs(title = "Scatterplot in ggplot2",
       subtitle = "Data source: Palmer Penguins",
       caption = "SML 201")
```

## Highcharter

We will create a simple scatterplot in `highcharter`.

```{r}
hchart(
  penguins,
  "scatter",
  hcaes(x = flipper_length_mm, y = bill_length_mm,
        group = species)
) |>
  hc_title(text = "Scatterplot in Highcharter") |>
  hc_subtitle(text = "Data source: Palmer Penguins") |>
  hc_caption(text = "SML 201")
```

