---
title: "Ghibli Palettes"
author: "Derek Sollberger"
date: "2024-01-27"
format: html
---

Today, inspired by [this article](https://medium.com/@jchen001/r-ggplot2-color-palettes-inspired-by-hayao-miyazakis-animes-f2aeccce45fd), I want to try out the `ghibli` package of color palettes inspired by some of the famous Studio Ghibli animated movies.

```{r}
library("ghibli")
library("ggplot2")
```

For some data, I was looking at the [demographics at Princeton University](https://inclusive.princeton.edu/about/demographics). I am to create a simple data frame with the proportion of Asian people in each of various populations found at the university.

```{r}
asian_percent <- c(28, 27, 33, 31, 20, 11, 11, 14, 11, 12)
populations <- c("undergraduates", "master's students", "doctoral students", "postdocs", "assistant professors", "associate professors", "full professors", "non-tenure-track faculty", "senior staff", "all other staff")

df <- data.frame(asian_percent, populations)
```

Here is a graph made with some of the `ggplot2` default settings.

```{r}
df |>
  ggplot(aes(x = asian_percent, y = populations, fill = asian_percent)) +
  geom_bar(stat = "identity")
```

And here is a graph made with a palette inspired by the movie *Kiki's Delivery Service*

```{r}
df |>
  ggplot(aes(x = asian_percent, y = populations, fill = asian_percent)) +
  geom_bar(show.legend = FALSE,
           stat = "identity") +
  labs(title = "Asian Populations at Princeton",
       subtitle = "Academic Year 2022-2023",
       x = "percent",
       y = "") +
  scale_fill_ghibli_c("KikiMedium") +
  theme_minimal()
```

In this quick exploration, it appears that the color scales can be applied only to continuous scales.