---
title: "Canvas_Roster"
author: "Derek Sollberger"
date: "2023-01-11"
format: html
---

With the goal of learning my students' names faster, I wrote an R script that will convert a roster ("People" in a Canvas course) to a `gt` table for later offline use.

```{r}
knitr::opts_chunk$set(eval = FALSE)
```


# Setup

In your Canvas course, go to `People`, right-click on the roster, and click "Save Page As"

* be sure to save the "complete" web page (this creates a folder with all of the students' pictures)
* place the also downloaded HTML file into the folder
* place this code script into the folder

```{r}
# change this file name as needed
HTML_file_name <- "Course Roster S23-BIO 018 01.htm"

library("gt")
library("gtExtras")
library("rvest")
library("tidyverse")
```

# Roster table

https://www.r-bloggers.com/2020/04/scrape-html-table-using-rvest/

```{r}
content <- read_html(HTML_file_name)
tables <- content |> html_table(fill = TRUE)
roster_table <- tables[[1]]

roster_df <- roster_table |>
  select(Name, Section) |>
  mutate(discussion_tag = str_extract(Section, "Discussion-\\d\\dD")) |>
  separate(discussion_tag, sep = "-", into = c("disregard", "discussion_section")) |>
  select(Name, discussion_section) |>
  
  # remove instructor of record
  filter(!is.na(discussion_section))
```

# Images

https://community.rstudio.com/t/scraping-images-from-the-web/133239

```{r}
image_urls <- content |> html_elements("img")
num_images <- length(image_urls)

image_df_raw <- data.frame(html_raw = rep(NA, num_images))
for(img in 1:num_images){
  image_df_raw$html_raw[img] <- as.character(image_urls[[img]])
}

image_df <- image_df_raw |>
  
  # skip first two images?
  slice(3:n()) |>
  
  # roundabout because some people's names have more than 2 words
  separate(col = "html_raw", sep = "alt=", 
           into = c("part1", "part2")) |>
  
  separate(col = "part1", sep = " ", into = c("tag1", "src")) |>
  separate(col = "src", sep = "/", into = c("stem", "profile_picture")) |>
  separate(col = "part2", sep = " aria", into = c("student_name", "tag2")) |>
  
  select(profile_picture, student_name)

# remove quotation marks
image_df$profile_picture <- str_replace_all(image_df$profile_picture, "\"", "")
image_df$student_name <- str_replace_all(image_df$student_name, "\"", "")
```

# Merge

```{r}
roster_df <- roster_df |>
  left_join(image_df, by = c("Name" = "student_name")) |>
  select(profile_picture, Name, discussion_section)
```

# gt

https://jthomasmock.github.io/gtExtras/reference/gt_img_rows.html

```{r}
roster_gt <- roster_df |>
  gt() |>
  gt_img_rows(columns = profile_picture, img_source = "local")
```

```{r}
gtsave(roster_gt, "roster.png")
```

