 ---
title: "Flint Water Crisis"
author: "Derek Sollberger"
date: "2026-01-24"
format:
  html:
    toc: true
---

# Tidy Tuesday

The TidyTuesday [data set](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-11-04/readme.md) for November 4, 2025, was about the Flint water crisis. I want to summarize some of the findings for my data science course.

```{r}
#| message: false
#| warning: false

library("tidyverse")

# load data (as shown)
flint_mdeq <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-11-04/flint_mdeq.csv')
flint_vt <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-11-04/flint_vt.csv')
```

# Summary statistics

## All MDEQ Samples

```{r}
summary(flint_mdeq$lead)
```

## Outliers Removed

```{r}
summary(flint_mdeq$lead2)
```

## Independent Test

```{r}
summary(flint_vt$lead)
```

```{r}
mu1_mdeq <- mean(flint_mdeq$lead, na.rm = TRUE)
mu2_mdeq <- mean(flint_mdeq$lead2, na.rm = TRUE)
mu_vt <- mean(flint_vt$lead, na.rm = TRUE)

top10_1_mdeq <- quantile(flint_mdeq$lead, 0.90, na.rm = TRUE)
top10_2_mdeq <- quantile(flint_mdeq$lead2, 0.90, na.rm = TRUE)
top10_vt <- quantile(flint_vt$lead, 0.90, na.rm = TRUE)
```


# Histograms

## All MDEQ Samples
\
```{r}
flint_mdeq |>
  ggplot() +
  geom_histogram(aes(x = lead),
                 binwidth = 5, color = "black", fill = "gray75") +
  geom_vline(xintercept = mu1_mdeq, color = "blue", 
             linetype = 2, linewidth = 2) +
  geom_vline(xintercept = top10_1_mdeq, color = "red", 
             linetype = 2, linewidth = 2) +
  labs(title = "Flint Water Samples in 2015",
       subtitle = "MDEQ measurements (average in blue, top 10 percentile in red)",
       caption = "Source: Jen Richmond",
       x = "lead (ppb)", y = "number of samples") +
  theme_minimal()
```

## Outliers Removed

```{r}
flint_mdeq |>
  ggplot() +
  geom_histogram(aes(x = lead2),
                 binwidth = 5, color = "black", fill = "gray75") +
  geom_vline(xintercept = mu1_mdeq, color = "blue", 
             linetype = 2, linewidth = 2) +
  geom_vline(xintercept = top10_2_mdeq, color = "red", 
             linetype = 2, linewidth = 2) +
  labs(title = "Flint Water Samples in 2015",
       subtitle = "MDEQ measurements, two samples missing (average in blue, top 10 percentile in red)",
       caption = "Source: Jen Richmond",
       x = "lead (ppb)", y = "number of samples") +
  theme_minimal()
```

## Independent Test

```{r}
flint_vt |>
  ggplot() +
  geom_histogram(aes(x = lead),
                 binwidth = 5, color = "black", fill = "gray75") +
  geom_vline(xintercept = mu1_mdeq, color = "blue", 
             linetype = 2, linewidth = 2) +
  geom_vline(xintercept = top10_vt, color = "red", 
             linetype = 2, linewidth = 2) +
  labs(title = "Flint Water Samples in 2015",
       subtitle = "VT measurements (average in blue, top 10 percentile in red)",
       caption = "Source: Jen Richmond",
       x = "lead (ppb)", y = "number of samples") +
  theme_minimal()
```

# Diagnoses

"the Lead and Copper Rule (LCR) of 1991 is 15 parts per billion (ppb). If this is exceeded in more than 10% of homes tested (or if the 90th percentile value of the total sample is above 15 ppb), action is required." --- [Significance, Vol 14, Issue 2](https://academic.oup.com/jrssig/article/14/2/16/7029247?login=false)

## All MDEQ Samples

```{r}
ifelse(quantile(flint_mdeq$lead, 0.90, na.rm = TRUE) > 15,
       "action required", "safe water")
```

## Outliers Removed

```{r}
ifelse(quantile(flint_mdeq$lead2, 0.90, na.rm = TRUE) > 15,
       "action required", "safe water")
```

## Independent Test

```{r}
ifelse(quantile(flint_vt$lead, 0.90, na.rm = TRUE) > 15,
       "action required", "safe water")
```


# Coda

::: {.callout-note collapse="true"}
## Session Info

```{r}
sessionInfo()
```
:::


:::: {.columns}

::: {.column width="45%"}
	
:::

::: {.column width="10%"}
	
:::

::: {.column width="45%"}

:::

::::

::::: {.panel-tabset}



:::::