---
title: "Andre Ethier versus Brad Lidge"
author: "Derek Sollberger"
date: "2025-04-25"
format:
  html
---

```{r}
#| message: false
#| warning: false

library("Lahman")
library("tidyverse")
```


# Setting the Scene

I am finally reading [Analyzing Baseball Data with R](https://beanumber.github.io/abdwr3e/), and thankfully there is a book club over at the DSLC Slack channel (Data Science Learning Community).

As we proceed through the first few chapters, I want to apply our new skills to a particular half-inning that is cherished in my memories. In particular, I want to see if I can find a particular at-bat by Andre Ethier where he faced Brad Lidge in the bottom of the ninth ...

## Lahman

The [Lahman](https://github.com/cdalzell/Lahman) data set can help us find overall information through its `People`, `Batting`, `Pitching`, `Fielding`, and `Teams` tables (and this data set is more extensive than I remember). First, let us find certain players in the `People` table:

```{r}
Lahman::People |>
  filter(nameLast %in% c("Ethier", "Lidge", "Werth"))
```

We will need these player IDs (the latter one in the case of "Werth"). I will now try to look up Ethier, Lidge, and Werth in the `Batting`, `Pitching`, and `Fielding` tables respectively.

::::: {.panel-tabset}

### Batting

```{r}
Lahman::Batting |>
  filter(playerID == "ethiean01")
```

### Pitching

```{r}
Lahman::Pitching |>
  filter(playerID == "lidgebr01")
```

### Fielding


```{r}
Lahman::Fielding |>
  filter(playerID == "werthja01")
```

:::::

::: {.callout-note collapse="true"}
## Session Info

```{r}
sessionInfo()
```
:::


:::: {.columns}

::: {.column width="45%"}
	
:::

::: {.column width="10%"}
	
:::

::: {.column width="45%"}

:::

::::

::::: {.panel-tabset}



:::::