---
title: "Andre Ethier versus Brad Lidge"
author: "Derek Sollberger"
date: "2025-07-16"
format:
  html
---

```{r}
#| message: false
#| warning: false

library("Lahman")
library("retrosheet")
library("tidyverse")
```


# Setting the Scene

I am finally reading [Analyzing Baseball Data with R](https://beanumber.github.io/abdwr3e/), and thankfully there is a book club over at the DSLC Slack channel (Data Science Learning Community). I want to apply our new skills to a particular half-inning that is cherished in my memories. In particular, I want to see if I can find *a particular at-bat by Andre Ethier where he faced Brad Lidge in the bottom of the ninth* ...

After volunteering to present the chapter on Quarto for our book club, I have decided to proceed through the materials as if I was giving a review session.


# Ch 1: Data Sets

## Lahman

The [Lahman](https://github.com/cdalzell/Lahman) data set can help us find overall information through its `People`, `Batting`, `Pitching`, `Fielding`, and `Teams` tables (and this data set is more extensive than I remember). First, let us find certain players in the `People` table:

```{r}
Lahman::People |>
  filter(nameLast %in% c("Ethier", "Lidge", "Werth"))
```

We will need these player IDs (the latter one in the case of "Werth"). I will now try to look up Ethier, Lidge, and Werth in the `Batting`, `Pitching`, and `Fielding` tables respectively.

::::: {.panel-tabset}

### Batting

```{r}
Lahman::Batting |>
  filter(playerID == "ethiean01")
```

### Pitching

```{r}
Lahman::Pitching |>
  filter(playerID == "lidgebr01")
```

### Fielding


```{r}
Lahman::Fielding |>
  filter(playerID == "werthja01")
```

:::::

## Retrosheet

Let us try out the [retrosheet package](https://cran.r-project.org/web/packages/retrosheet/index.html) (by Colin Douglas and Richard Scriven) to get **game data** from each of those four seasons where *Brad Lidge was pitching for the Philadelphia Phillies*. Note: we probably do not have to concern ourselves with mid-season trades in this case study.

```{r}
season_2008 <- retrosheet::get_retrosheet("game", 2008)
season_2009 <- retrosheet::get_retrosheet("game", 2009)
season_2010 <- retrosheet::get_retrosheet("game", 2010)
season_2011 <- retrosheet::get_retrosheet("game", 2011)
```


## Statcast

As recommended by the textbook authors in section C.2, we can load a list of player IDs for Statcast data (takes a few seconds to load and download)

```{r, eval = FALSE}
chadwick_id <- baseballr::chadwick_player_lu()
readr::write_rds(chadwick_id, "data/chadwick_id.rds")
chadwick_id <- readr::read_rds("data/chadwick_id.rds")

statcast_ids <- chadwick_id |>
  filter(name_last %in% c("Ethier", "Lidge", "Werth")) |>
  select(key_mlbam, name_last, name_first, mlb_played_first, mlb_played_last) |>
  filter(!is.na(mlb_played_first))
readr::write_rds(statcast_ids, "data/statcast_ids.rds")
```

```{r}
statcast_ids <- readr::read_rds("data/statcast_ids.rds")
statcast_ids
```


# Ch 2: dplyr

## Retrosheet

For this case study, we can use the `filter` command from the `dplyr` to subset our data where

* visiting team: Philadelphia Phillies
* home team: Los Angeles Dodgers

```{r}
lidge_phillies_dodgers <- rbind(season_2008, season_2009, season_2010, season_2011) |>
  dplyr::filter(VisTm == "PHI" & HmTm == "LAN")
lidge_phillies_dodgers |> select(Date, VisTm, HmTm, VisRuns, HmRuns, Attendance)
```

## Statcast

In addition to the calendar dates found above, we can try to get pitch-by-pitch data from the `statcast_search` function in the `baseballr` package.

```{r, eval = FALSE}
lidge1 <- baseballr::statcast_search(start_date = "2008-08-11", end_date = "2008-08-11", playerid = 400058)
```

[Derek forgot that Statcast data only dates back to 2015]



















::: {.callout-note collapse="true"}
## Session Info

```{r}
sessionInfo()
```
:::


:::: {.columns}

::: {.column width="45%"}
	
:::

::: {.column width="10%"}
	
:::

::: {.column width="45%"}

:::

::::

::::: {.panel-tabset}



:::::