---
title: "California Weather"
author: "Derek Sollberger"
date: "2023-01-25"
format: html
---

```{r}
#| message: false
#| warning: false
library("tidyverse")
```

I want to create and visualize a simple data set for my Data Science courses (that I teach in California).

# Data Source

* University of California
* Agriculture and Natural Resources
* Statewide Integrated Pest Management Program
* https://ipm.ucanr.edu/WEATHER/wxactstnames.html

# Fixed-Width Files

Today I learned how to read `fixed-width files` in the [Tidyverse](https://readr.tidyverse.org/).  From there, I simply need to give the columns easy-to-use names.

```{r}
LA_df <- readr::read_fwf("LA_2022.txt")
colnames(LA_df) <- c("date", "time", "precipitation",
                     "check1", "high", "low", "check2", "solar", "check3")
LA_df$city <- "Los Angeles"
```

```{r}
Merced_df <- readr::read_fwf("Merced_2022.txt")
colnames(Merced_df) <- c("date", "time", "precipitation",
                         "check1", "high", "low", "check2")
Merced_df$city <- "Merced"
```

```{r}
SF_df <- readr::read_fwf("SF_2022.txt")
colnames(SF_df) <- c("date", "time", "precipitation",
                     "check1", "high", "low", "check2")
SF_df$city <- "San Francisco"
```

# Merge

Some of the weather stations had collected more information than others.  That is, if the weather station was newer, then it had more instruments.

For today's quick exploration, I actually do want to perform a quick `rbind`, and that requires that each of the 3 data frames have the same number of columns (and should be the same types of information too).

```{r}
LA_df <- LA_df |>
  select(city, date, time, high, low, precipitation)
Merced_df <- Merced_df |>
  select(city, date, time, high, low, precipitation)
SF_df <- SF_df |>
  select(city, date, time, high, low, precipitation)

CA_weather_data <- rbind(LA_df, Merced_df, SF_df)
```

```{r}
# write_csv(CA_weather_data, "CA_weather_data.csv")
```


# Data Viz

Now, boxplots are easy to make.

```{r}
CA_weather_data |>
  ggplot(aes(y = high)) +
  geom_boxplot() +
  labs(title = "California Weather, High Temperatures",
       subtitle = "(all together)",
       caption = "Source: UC\nAgriculture and Natural Resources\nStatewide Integrated Pest Management Program")
```

```{r}
CA_weather_data |>
  ggplot(aes(x = city, y = high, fill = city)) +
  geom_boxplot() +
  labs(title = "California Weather, High Temperatures",
       subtitle = "(separate groups)",
       caption = "Source: UC\nAgriculture and Natural Resources\nStatewide Integrated Pest Management Program")
```

# Sample

For the creation of a classroom example, I want to randomly select 43 observations from the Merced data.

```{r}
Merced_sample <- sort(sample(Merced_df$high, 43, replace = FALSE))
dput(Merced_sample)
```

