---
title: "ggflowchart"
author: "Derek Sollberger"
date: "2024-05-18"
format: html
---

Today I explore the `ggflowchart` [package](https://nrennie.rbind.io/blog/introducing-ggflowchart/) (made by Dr Nicola Rennie).  For some data, the scenario comes from [The Planet Crafter](https://mijugames.com/pages/planetcrafter/presskit.html) video game.  Later in the game (minor spoiler), you can fire rockets into space to *deliver* and trade items for currency.  Before then, these items are *crafted* out of materials that are *gathered* on the planet.  The information comes from [The Planet Crafter wiki page](https://planet-crafter.fandom.com/wiki/Planet_Crafter_Wiki).  This blog post is not meant to be an accurate guide for gameplay; I took some liberties (and also did not indicate the quantities of the ingredients) in this quick exploration.

```{r}
#| message: false
#| warning: false
library("ggflowchart")
library("readxl")
```


# Data

I transcribed information for the nodes and edges into an Excel spreadsheet for typing convenience.

:::: {.panel-tabset}

## Loading

```{r}
edge_df <- readxl::read_xlsx("planet_crafter_trade_goods.xlsx", sheet = "edges")
node_df <- readxl::read_xlsx("planet_crafter_trade_goods.xlsx", sheet = "nodes")
```

## Nodes

```{r}
node_df
```

## Edges

```{r}
edge_df
```

::::

## Network

Now let's see if the main `ggflowchart` function will work well with this situation.

```{r}
ggflowchart::ggflowchart(edge_df, node_df, fill = stage)
```
The code package did the job as prescribed correctly and quickly, but I had presented a situation that was slightly complicated for a default graph.

## Customization

Next, we try to adjust the network presentation with some of the underlying tools in the `igraph` package.

```{r}
ggflowchart(edge_df, node_df, fill = stage, horizontal = TRUE)
```

This graph is more readable.  From here, one could explicity define where the nodes are (by x and y coordinates).

In the game's jargon, it might have made more sense to group the materials by tiers ("T1", "T2", "T3") and perhaps repeat nodes that are used in multiple crafting recipes.